Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-27 Manuscript.Rmd
working on Joe's comments

---
title: 'Individual heterogeneity in the functional topography of the default-mode network in medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: 
    keep_tex: true
  word_document: 
    reference_docx: Style_reference_draft.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a large set of disparate cognitive functions [@Hiser2018; @Kragel2018], including the subjective value (SV) assigned to decision outcomes [@Bartra2013; @Clithero2014; @Hiser2018; @Kable2007; @Levy2011]. Choice-related activity in these areas has shown significant spatial overlap with the default mode network (DMN) along the medial wall [@Fox2005; @Laird2009; @Yeo2011], prompting the question of whether (and to what extent) these two systems can be dissociated.

  Previous coordinate-based meta-analyses have identified topographically similar activation patterns in mPFC for valuation [@Bartra2013; @Clithero2014] and default-mode task deactivation [@Laird2009]. More focused meta-analytic work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC [@Acikalin2017], even though there is clear divergence in other areas (ventral striatum in favor of SV, with DMN exclusively on sections of PCC and the temporoparietal junction). These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same neural system.

  A key limitation of group-averaging and meta-analysis is that individual-specific functionl subdivisions can be overlooked. Averaging has traditionally been beneficial in identifying tendencies in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding [@Zilles2013; @Mackey2014] and inter-subject functional variability [@Mueller2013a], thus making group-averaging studies difficult to interpret.

  Seeking to remedy these shortcomings, recent work has prescribed relevance to the analysis of single subjects in fMRI. Individual analyses of fMRI data have identified idiosyncratic, reliable, and valid functional organization that would otherwise be blurred in aggregative estimations [@Laumann2015; @Gordon2017; @Gratton2018a; @Tobyne2018]. Furthermore, subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. In regards to DMN, the trend of individualized analyses has led to finer idiosyncratic subdivisions of the DMN through careful selection of seed-based functional connectivity [@Braga2017]. It is thus possible that the indistinguishable overlap of DMN and valuation effects can be attributed to a lack of spatial resolution that is better understood at the individual level. With this in mind, an important first step in disentangling these phenomena would be to determine the degree of topological heterogeneity of the DMN present within mPFC.

  Interrogating all sources of covariation through functional connectivity provides an avenue for capturing subject-specific brain organization. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity [@Gordon2017; @Laumann2015; @Tobyne2017]. A popular approach to analyze these connectomes is to rely on graph theoretic methods, which are properly equiped to understand brain network dynamics [@Bassett2018a]. From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures [@Garcia2017]. Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network [@Fortunato2016]. Among the algorithms used to estimate these communities, modularity has been widely effective in subsectioning brain networks into multiple groups [@Garcia2017]. However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes a forced dissection through spectral partitioning (SP) a more valuable option [@Fiedler1975; @Belkin2003]. SP has recently been used to characterize the posterior-anterior functional gradient of the insula using rsfMRI data [@Tian2018]. This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

@@

We evaluated the performance of the spectral partitioning algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average PCC activity for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by spatially correlating them across days and subjects in three ways: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were aimed to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods were expected to produce somewhat similar arrangements, but the one displaying the higher within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

- Summarize the results
- Ours are not the first to show variability, but previous studies have focused on a relatively small number of subjects and often rely on qualitative assessment of these differences. They also use parcellations for networks, whereas we focused on individual vertices to gain spatial finesse. Also, to the best of our knowledge we are the first to quantify the organizational heterogeneity of these areas in the human brain. 
- We are also not the first to show functional subdivision of DMN [@Kernbach2018] (but that's using group averages on parcellated data), or this particular individualized pattern of subdivision (braga). Our results are complementary in the sense that we show similar outcomes in 100 subjects. The major discrepancy is that we assign meaning to the communities based on their coverage of regions identified through a meta-analysis.
- Talk about SP as a way to identify an individual's partition: deterministic, with simple diagnostics, and requires less data requirements.
- The main gap is mapping behavior to functional organization, but we will do that with behavioral manipulations of subjective value in the future. Instead, we focus on understanding the importance of individualized mapping in these areas first, as a baseline.
- Clinical implications: individualized targets, and more accurate estimation of impacts based on assessment of individual patients. Similarly, our approach focuses on individual estimations without group priors, so it's well fit to address the problem of low n in clinical research. Also, SP can produce these effects with moderate amounts of data. Previous results suggest impractical amounts.

>>>>

---
title: 'Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: 
    keep_tex: true
  word_document: 
    reference_docx: Style_reference_draft.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a large set of disparate cognitive functions [@Hiser2018; @Kragel2018]. One such function is subjective valuation during economic decision making [@Bartra2013; @Clithero2014; @Hiser2018; @Kable2007; @Levy2011]. Both mPFC and PCC (together with ventral striatum) consistently show greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued. These regions are also implicated in a set of functions associated with the default-mode network (DMN). DMN effects include a decrease in BOLD activity during cognitively demanding tasks compared with less-demanding task conditions or periods of rest [@McKiernan2003; @Laird2009], as well as a distinctive pattern of inter-region correlations in resting-state fMRI data [@Fox2005; @Yeo2011].

  Coordinate-based meta-analyses of the fMRI literature have demonstrated that valuation and DMN effects occur in overlapping and largely indistinguishable regions of ventral mPFC [@Acikalin2017]. The two sets of effects diverge in other brain areas; a region of the striatum is preferentially associated with valuation, whereas bilateral temporoparietal junction is linked with the DMN. Within PCC and adjacent precuneus, a more posterior subregion appears DMN-specific, whereas a more anterior subregion is implicated in both effects. One possible interpretation of these findings is that mPFC and anterior PCC operate as common nodes within the distributed brain systems that subserve valuation and DMN effects. The potential existence of common nodes has primpted theorizing about posible cognitive-level overlap between valuation and DMN-related processes. For example, the two sets of processes might share elements of self-referential cognition, episodic memory, mental simulation, or monitoring and regulation of internal states [@Northoff2012; @Clithero2014; @Acikalin2017; Reddan2018].

  However, to draw firm conclusions about functional overlap will require a better understanding of individual-level heterogeneity in topographic patterns of brain activity. A key limitation of group averaging and meta-analysis is that individual-specific functional subdivisions can be blurred and overlooked [@Michalka2015; @Fedorenko2012; Woo2014], potentially exaggerating the degree of overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Zilles2013; @Mackey2014] and inter-subject functional variability [@Mueller2013a]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Laumann2015; @Gordon2017; @Gratton2018a; @Tobyne2018]. Subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DMN through tailored analyses of seed-based functional connectivity [@Braga2017]. It is therefore possible that the diffuse overlap of DMN and valuation effects can be attributed to a lack of spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of DMN effects within mPFC and PCC.

  A useful way to characterize individual-specific brain organization is to examine patterns of resting-state functional connectivity (rsFC; i.e. time-series correlations between pairs of locations in the brain). Connectome-based analyses of rsFC have been fruitful in characterizing individualized functional topologies that correspond well to task-induced activity [@Gordon2017; @Laumann2015; @Tobyne2017; @Smith2009]. A popular way to analyze these connectomes is to rely on graph theoretic methods, which are properly equiped to understand brain network dynamics [@Bassett2018a]. In the context of network analysis, community detection algorithms have been successfully used to subdivide brain networks into multiple distinct communities [@Garcia2017]. Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network [@Fortunato2016]. Among these algorithms, spectral partitioning (SP) efficiently captures  [@Fiedler1975; @Belkin2003]. SP has recently been used to characterize the posterior-anterior functional gradient of the insula using rsfMRI data [@Tian2018]. This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

@@

We evaluated the performance of the spectral partitioning algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average PCC activity for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by spatially correlating them across days and subjects in three ways: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were aimed to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods were expected to produce somewhat similar arrangements, but the one displaying the higher within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Gratton2018a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

- Summarize the results
- Ours are not the first to show variability across individuals, but previous studies have focused on a relatively small number of subjects and often rely on qualitative assessment of these differences. They also use parcellations for networks, whereas we focused on individual vertices to gain spatial finesse. Also, to the best of our knowledge we are the first to quantify the organizational heterogeneity of these areas in the human brain. 
- We are also not the first to show functional subdivision of DMN [@Kernbach2018] (but that's using group averages on parcellated data), or this particular individualized pattern of subdivision (braga). Our results are complementary in the sense that we show similar outcomes in 100 subjects. The major discrepancy is that we assign meaning to the communities based on their coverage of regions identified through a meta-analysis.
- Stability within subject has also been showsn before (Gratton2018b). This work was done using parcellations and oversampling though.
- Talk about SP as a way to identify an individual's partition: deterministic, with simple diagnostics, and requires less data requirements.
- The main gap is mapping behavior to functional organization, but we will do that with behavioral manipulations of subjective value in the future. Instead, we focus on understanding the importance of individualized mapping in these areas first, as a baseline.
- Clinical implications: individualized targets, and more accurate estimation of impacts based on assessment of individual patients. Similarly, our approach focuses on individual estimations without group priors, so it's well fit to address the problem of low n in clinical research. Also, SP can produce these effects with moderate amounts of data. Previous results suggest impractical amounts.
